---
title: "Survival Analysis"
author: "Batoul Kalot"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(survival)
library(survminer)
library(dplyr)
library(readxl)
library(tidyverse)
library(tidymodels)
library(gt)
library(lsr)
library(survival)
library(rpart)
library(rpart.plot)
library(epiR)

```

## Loading data
```{r echo=FALSE}
colon <- read_excel("C:/Users/Lenovo/Downloads/colon.xlsx")
head(colon) %>% gt() %>% 
  opt_stylize(style=2,color="blue")
```

## Exploratory Analysis
```{r}
colon %>% 
  ggplot(aes(x=age))+
  geom_histogram(color="red")

```

**There exists some outliers under age 30**

## Filtering and preparing data
```{r}
data <- colon %>% 
  filter(age>30)
data <- data %>% 
         mutate(Time=as.Date(time))
```

## EDA
```{r}
data %>% ggplot(aes(x=status))+
                geom_bar(fill="green")
```

**Same count for colon status** 

```{r}
data %>% ggplot(aes(x=surg))+
  geom_bar(fill="yellow")

```

**Count of cases which are exposed to colon surgeries are much less than not**

## Checking relations and associations
```{r}
attach(data)
tab <- table(status, surg)
chi <- chisq.test(tab)
```


```{r}
print(chi)
```


```{r}
cramersV(tab)
```
**High effect size between surg and status of colon**

```{r}
attach(data)
tab2 <- table(status, sex)
chi <- chisq.test(tab2, correct=T)
```


```{r}
print(chi)
```
**Not significant p_value , so there is no relation between sex and status.**

```{r}
attach(data)
tab3 <- table(status,obstruct)
chi <- chisq.test(tab3, correct=T)
```

```{r}
print(chi)
```
**Not significant p_value.**

```{r}
attach(data)
tab4 <- table(status,adhere)
chi <- chisq.test(tab4, correct=T)
```

```{r}
print(chi)
```

```{r}
library(lsr)
cramersV(tab4)
```
**High effect size between adhesions and status of the colon**


## Time series analysis
```{r}
data %>%
  mutate(sex=as.factor(sex)) %>% 
  ggplot(aes(x=Time,
             y=nodes,
             color=sex))+
  geom_line()
```

# Survival Analysis
## Manipulating data 
```{r}
data <- data %>% 
  mutate(exposed = (nodes>12),
         status=as.factor(status),
         adhere=as.factor(adhere),
         surg=as.factor(surg),
         node4=as.factor(node4),
         Time=as.integer(Time)) %>% 
  select(Time, status, nodes, 
         adhere, surg, node4, exposed)
```

```{r}
head(data) %>% gt()
```

# Survival Analysis
## Creating our survival function
```{r}
colonsurv <- survfit(Surv(Time, node4) ~ exposed + status, data)
```


```{r}
plot(colonsurv, xlab="Time", ylab="Survival Probability",
     main="Kaplan-Meier Curve")
```

```{r}
cox.mode <- coxph(Surv(time,node4) ~ status)
```

```{r}
summary(cox.mode)
```
**So based on th hazard ratio in our cox model, The group with positive status of colon exposed to node4 has a risk of the cancer that is 20.4 times than the group with status 0. **


# Logistic Regression
```{r}
set.seed(2)
 split <- initial_split(data,
                        prop=.80,
                        strata=node4)  
 data_train <- training(split) 
 data_test <- testing(split)
```

# Building the model
```{r}
model <- glm(node4 ~ status,
              data=data_train,
              family="binomial")
```


```{r}
summary(model)
```

## Testing the accuracy
```{r}
data_test <- data_test %>% 
   mutate(node4_prob = predict(model,
                                data_test,
                                type= "response"),
          node4_pred = ifelse(node4_prob > .5, 1, 0))
t <- table(data_test$node4,
            data_test$node4_pred)
```


```{r}
accuracy <- sum(diag(t))/sum(t)     

```

```{r}
print(accuracy)
```
**73% accuracy of the classification model built based on status of the colon**

### Decision tree based on exposed cases(have nodes>12)
```{r}
tree_model <- rpart(node4 ~ exposed,data=data_train,
                    method = "class")
rpart.plot(tree_model, main = "Decision Tree for colon Classification")
```

# Incidence: Risk and Odds Ratios according to the surgery
```{r}
TAB <- table(surg, node4)
epi.2by2(TAB, method="cohort.count",
         conf.level=0.95)
```
**96% risk ratio and 86% odds ratio, means that sample people who made the surgery had 96% of the risk of reaching node4 level for colon compared to the non_surgeried people sample.**

**This would suggest that the surgery is providing some protection,** 
**but the effect is small.**


